---
title: "Measures_demo"
author: "Brian Schilder"
date: "March 20, 2015"
output: html_document
---

# Measures demo for brain-climate poster: Trend, Variability, and Rate

## Trend
```{r}
library(ggplot2)

Trend <- data.frame(c(1, 1), c(2.5, 2.5), c(0, 1), c(1, 2))

Trend_low <- ggplot() + geom_line(aes(x=Trend[,3], y=Trend[,1]), col="blue", size=2) + labs(x=" ", y="δ18O", title="Low") + 
  theme_classic(40) +  scale_x_continuous(breaks = round(seq(0, 1, by = 0.5),1)) + ylim(0,3) + theme(plot.title = element_text(vjust=3))
Trend_low

Trend_high <- ggplot() + geom_line(aes(x=Trend[,4], y=Trend[,2]), col="blue", size=2) + labs(x=" ", y=" ", title="High") + 
  theme_classic(40) +  scale_x_continuous(breaks = round(seq(1, 2, by = 0.5),1)) + 
  scale_y_continuous(breaks = round(seq(0, 3, by = 1),1)) + ylim(0,3) + theme(plot.title = element_text(vjust=3))
Trend_high 

```

## Variability

```{r}
x <- seq(0, 1, 0.01)
ylow <- sin(10*x)
yhigh <- sin(50*x)
Var <- data.frame(x, ylow, yhigh)

Var_low <- ggplot() + geom_line(aes(x=Var$x, y=Var$ylow), col="red", size=2) + labs(x=" ", y="δ18O") + theme_classic(40) + scale_x_continuous(breaks = round(seq(0, 1, by = 0.5),1)) + scale_y_continuous(breaks = round(seq(-1, 1, by = 1),1))
Var_low

Var_high <- ggplot() + geom_line(aes(x=Var$x, y=Var$yhigh), col="red", size=2) + labs(x=" ", y=" ") + theme_classic(40) + scale_x_continuous(breaks = round(seq(0, 1, by = 0.5),1)) + scale_y_continuous(breaks = round(seq(-1, 1, by = 1),1))
Var_high
```

## Rate

```{r}

Rate_low <- ggplot() + geom_line(aes(x=Trend[,3], y=Trend[,1]), col="white") +
  geom_abline(intercept=0.5, slope=0.25, col="green", size=2) + 
  labs(x=" ", y="δ18O") + theme_classic(40) +  
  scale_x_continuous(breaks = round(seq(0, 1, by = 0.5),1)) + scale_y_continuous(breaks = round(seq(0, 2, by = .5),1)) 
Rate_low 

Rate_high <- ggplot() + geom_line(aes(x=Trend[,3], y=Trend[,1]), col="white") +
  geom_abline(intercept=0.5, slope=1, col="green", size=2) + labs(x=" ", y=" ") + 
  theme_classic(40) +  
  scale_x_continuous(breaks = round(seq(0, 1, by = 0.5),1)) + scale_y_continuous(breaks = round(seq(0, 2, by = .5),1))
Rate_high 

```

## Multiplot

```{r}
# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

## Combined plot

```{r fig.width = 2, fig.height = 2}
multiplot(Trend_low, Var_low, Rate_low, Trend_high, Var_high, Rate_high, cols=2)
```

# Interval Demo

```{r}
detach(package:dplyr, unload = TRUE)
library(scales)
zachos <- read.table("Zachos2001Climate.txt", header=TRUE, sep="\t")
zachos$filteredO18 <- filter(zachos$o18, filter = rep(1, 70), method = "convolution")

hominin <- read.table("Hominin_climatebrains_means.csv", header=TRUE, sep=",")
hominin$Species <- gsub(" ", "_", hominin$Species)
hominin$MYA <- abs(hominin$MYA)

plot <- ggplot(data=fossils, aes(color=Order))

Demo <- plot +
  geom_rect(aes(xmin=4.4, xmax=5.4, ymin=0, ymax=8), fill="grey") +
  geom_line(data=zachos,aes(x=age*1, y=rescale(filteredO18*1,c(-1,8))), color="black") +
  geom_point(data=hominin, aes(x=MYA, y=log(CC), fill=Order), size=6) +
  scale_x_reverse() +
  xlim(10, 0) +
  labs(x="MYA", y="log(CC)", title="Measurement Example") +
  theme_bw() +
  theme(text = element_text(size=40), plot.title = element_text(vjust=3), legend.position="none")
Demo
