# New analysis with Shultz-Dunbar mammals plus Boddy Primates

### Load datasets and clean up species columns and column names
```{r}
library(dplyr)
brains <- read.table("brains2.txt", header=TRUE, sep="\t")
#brains <- filter(brains, Order != "Insectivore") %>% droplevels() ## All brains except Insectivores
# brains <- filter(brains, Order == "Artiodactyla") %>% droplevels()
 brains <- filter(brains, Order == "Carnivora") %>% droplevels()
# brains <- filter(brains, Order == "Cetacea") %>% droplevels()
# brains <- filter(brains, Order == "Perissodactyla") %>% droplevels()
# brains <- filter(brains, Order == "Primates") %>% droplevels()
```

## Redo ancestral state with composite dataset 

```{r cache=TRUE}
library(ape)
library(geiger)
library(xlsx)
library(ggplot2)
treez <- read.nexus("Bininda_Emonds_correctedDates_FULLPHYLOG.txt")
tree <- treez[[grep(pattern = "bestDates", names(treez))]] #get the best dates tree
rm(treez)

name_matchup <- name.check(phy = tree, data = brains, data.names = brains$Species)
tree <- drop.tip(phy = tree, name_matchup$tree_not_data)
tree <- multi2di(tree)
tree$edge.length[tree$edge.length == 0] <- 0.1
tree <- makeLabel(tree) #Fills in empty node names
brains <- brains[brains$Species %in% tree$tip.label,]

stopifnot(nrow(brains)==length(tree$tip.label))

brains_CC <- brains$CC
names(brains_CC) <- brains$Species

ancest <- ace(x = brains_CC, phy = tree, type="continuous", method="REML")

```

## Export plot 

```{r}

#pdf("CompositeAncestral.pdf", width=20, height=200)
#plot(tree)
#nodelabels(round(ancest$ace, 2), cex=1.2)
#tiplabels(round(brains$CC, 2))
#dev.off()

#plot.phylo(tree, cex=0.7)
#tiplabels(text = round(brains$CC, 2)



#library(xlsx)
#Export a file containing species name (or ancestral node #) with associated CC
#Output <- c(brains_CC, ancest$ace)
#write.xlsx(Output, "Export.xlsx")

```

# Calculate Climatic Measures

## Prep Data

```{r}
library(ape)
zachos <- read.table("Zachos2001Climate.txt", header=TRUE, sep="\t")
ancest_climate <- numeric(326)

names(ancest$ace) <- tree$node.label
str(ancest$ace)

#Calculate time from all node to tip
mya <- branching.times(phy = tree)
nodes <- names(mya)
fdata <- data.frame(mya, nodes, CC = log(ancest$ace), orig_CC =ancest$ace) # Log and then scale CC 
rownames(fdata) <- NULL
branch65 <- dplyr::filter(fdata, mya <= 65)

data <- branch65$mya

results <- numeric(254)

plot(tree)
nodelabels(text = tree$node.label, frame="none", cex=0.75)

# Identifies most recent common ancestor
  ## mrca(phylogeny)["taxon1", "taxon2"]
```

## 1MY Measures

```{r}
library(dplyr)
## MEAN
Mean1MY <- function(data){
  start.time <- data
  end.time <- start.time + 1
  filteredZachos <- filter(zachos, 
                         age >= start.time, 
                         age <= end.time)
  results <- mean(filteredZachos$o18)
  return(results) }
mean1MY <- sapply(X=data, FUN=Mean1MY)
branch65$mean1MY <- scale(mean1MY) #scale at the end


## SD
SD1MY <- function(data){
  start.time <- data
  end.time <- start.time + 1
  filteredZachos <- filter(zachos, 
                         age >= start.time, 
                         age <= end.time)
  results <- sd(filteredZachos$o18)
  return(results) }
sd1MY <- sapply(X=data, FUN=SD1MY)
branch65$sd1MY <- scale(sd1MY)


## SLOPE
Slope1MY <- function(data){
  start.time <- data
  end.time <- start.time + 1
  filteredZachos <- filter(zachos, 
                         age >= start.time, 
                         age <= end.time)
  Model <- lm(o18 ~ age, data=filteredZachos)
  slope <- Model$coefficients[2]
  results <- slope
  return(results) }
slope1MY <- sapply(X=data, FUN=Slope1MY)
branch65$slope1MY <- scale(slope1MY)
```

## 400KY Measures

```{r}
## MEAN
Mean400KY <- function(data){
  start.time <- data
  end.time <- start.time + .4
  filteredZachos <- filter(zachos, 
                         age >= start.time, 
                         age <= end.time)
  results <- mean(filteredZachos$o18)
  return(results) }
mean400KY <- sapply(X=data, FUN=Mean400KY)
branch65$mean400KY <- scale(mean400KY)


## SD
SD400KY <- function(data){
  start.time <- data
  end.time <- start.time + .4
  filteredZachos <- filter(zachos, 
                         age >= start.time, 
                         age <= end.time)
  results <- sd(filteredZachos$o18)
  return(results) }
sd400KY <- sapply(X=data, FUN=SD400KY)
branch65$sd400KY <- scale(sd400KY)


## SLOPE
Slope400KY <- function(data){
  start.time <- data
  end.time <- start.time + .4
  filteredZachos <- filter(zachos, 
                         age >= start.time, 
                         age <= end.time)
  Model <- lm(o18 ~ age, data=filteredZachos)
  slope <- Model$coefficients[2]
  results <- slope
  return(results) }
slope400KY <- sapply(X=data, FUN=Slope400KY)
branch65$slope400KY <- scale(slope400KY)
```

## 200KY Measures

```{r}
## MEAN
Mean200KY <- function(data){
  start.time <- data
  end.time <- start.time + .2
  filteredZachos <- filter(zachos, 
                         age >= start.time, 
                         age <= end.time)
  results <- mean(filteredZachos$o18)
  return(results) }
mean200KY <- sapply(X=data, FUN=Mean200KY)
branch65$mean200KY <- scale(mean200KY)


## SD
SD200KY <- function(data){
  start.time <- data
  end.time <- start.time + .2
  filteredZachos <- filter(zachos, 
                         age >= start.time, 
                         age <= end.time)
  results <- sd(filteredZachos$o18)
  return(results) }
sd200KY <- sapply(X=data, FUN=SD200KY)
branch65$sd200KY <- scale(sd200KY)


## SLOPE
Slope200KY <- function(data){
  start.time <- data
  end.time <- start.time + .2
  filteredZachos <- filter(zachos, 
                         age >= start.time, 
                         age <= end.time)
  Model <- lm(o18 ~ age, data=filteredZachos)
  slope <- Model$coefficients[2]
  results <- slope
  return(results) }
slope200KY <- sapply(X=data, FUN=Slope200KY)
branch65$slope200KY <- scale(slope200KY)
```

#ALL ORDERS Multiple Regressions

```{r}
library(pander)
##ALL ORDERS: 1Mya
ALLsparse_results_1MY <- list(model = lm(branch65$CC ~ branch65$mean1MY + branch65$sd1MY + branch65$slope1MY))
pander(ALLsparse_results_1MY$model)

##ALL ORDERS: 400K
ALLsparse_results_400KY <- list(model = lm(branch65$CC ~ branch65$mean400KY + branch65$sd400KY + branch65$slope400KY))
pander(ALLsparse_results_400KY$model) 

##ALL ORDERS: 200K
ALLsparse_results_200KY <- list(model = lm(branch65$CC ~ branch65$mean200KY + branch65$sd200KY + branch65$slope200KY))
pander(ALLsparse_results_200KY$model)
```

# PLOTS

## Extant CC vs. Time

```{r}
library(dplyr)
library(pander)

  ## Separate Extants and Fossils
exbrains <-read.table(file="brainclimatedata.txt", header=TRUE, sep="\t") ## Old data
  exbrains$Actual.MYA <- exbrains$Actual.MYA * -1
  exbrains$MYA <- exbrains$MYA * -1
extants <- dplyr::filter(exbrains, Fossil == "Yes")
fossils <- dplyr::filter(exbrains, Fossil == "No")

  ## Extant CC over time
model3 <- lm(data=filter(.data=extants, Order != "Insectivore"), log(CC) ~ Actual.MYA)
  plot(extants$Actual.MYA, log(extants$CC), main = "All Orders")
  abline(model3)
  pander(model3)
```

## Fossil CC over time

```{r}
  ## All Orders
model4 <- lm(data=filter(.data=fossils, Order != "Insectivore"), log(CC) ~ Actual.MYA)
  plot(fossils$Actual.MYA, log(fossils$CC))
  abline(model4)
  pander(model4)
  
 ## Artiodactyla
model4 <- lm(data=filter(.data=fossils, Order == "Artiodactyla"), log(CC) ~ Actual.MYA)
  plot(fossils$Actual.MYA, log(fossils$CC))
  abline(model4)
  pander(model4)

## Carnivora
model4 <- lm(data=filter(.data=fossils, Order == "Carnivora"), log(CC) ~ Actual.MYA)
  plot(fossils$Actual.MYA, log(fossils$CC))
  abline(model4)
  pander(model4)

## Cetacea
model4 <- lm(data=filter(.data=fossils, Order == "Cetacea"), log(CC) ~ Actual.MYA)
  plot(fossils$Actual.MYA, log(fossils$CC))
  abline(model4)
  pander(model4)
```

# Ancestral CC over time

```{r}
 ## Filter order at beginning of script
model5 <- lm(data=branch65, CC ~ mya)
  plot(branch65$mya, branch65$CC)
  abline(model5)
  pander(model5)

## Whether it's raw, logged, logged+scaled; ancestral CCs are always negatively correlated with time. This makes sense since as time gets smaller (65-0) brains get bigger. Had opposite valence in previous analyses bc time was negative.
```

## Fancy plots

```{r}
detach(package:dplyr, unload = TRUE)
library(scales)
zachos <- read.table("Zachos2001Climate.txt", header=TRUE, sep="\t")
zachos$filteredO18 <- filter(zachos$o18, filter = rep(1, 70), method = "convolution")
zachos$filteredO18 <- as.numeric(zachos$filteredO18)


library(ggplot2)
plot <- ggplot(data=branch65)

## Climatic Variables
climate <- plot + geom_line(aes(x=mya, y=mean1MY), color="blue") +
  geom_line(aes(x=mya, y=sd1MY), color="red") +
  geom_line(aes(x=mya, y=slope1MY), color="green") +
  scale_x_reverse()
climate

#Extant CCs Plot
Extant_plot <- plot +
  geom_line(data=zachos, aes(x=age, y=rescale(filteredO18, c(-1,8))), color="blue") +
  scale_x_reverse() +
  geom_point(data=extants , aes(x=Actual.MYA, y=log(CC), color=Order), size=4) +
  labs(x="MYA", y="log(CC)", title="Extant CC Over Time") +
  theme(text = element_text(size=40)) +
  geom_smooth(data=extants, aes(x=Actual.MYA, y=log(CC)), color="red", method=lm, se=FALSE, inherit.aes=FALSE)
Extant_plot

# Extant CCs as facets
Facet_order_plot <- plot + 
  geom_line(data=zachos,aes(x=age, y=rescale(filteredO18,c(-1,8))), color="blue") +
  geom_point(data=exbrains, aes(y=log(CC), x=Actual.MYA, color=Order), size=3) + 
  geom_smooth(data=exbrains, aes(Actual.MYA, log(CC)), method="lm", color="red") +
  scale_x_reverse() +
  facet_wrap(~Order) +
  theme_bw(40)  +
  theme(plot.title = element_text(vjust=3), legend.position="none", strip.background = element_rect(fill="black"), strip.text.x = element_text(color="white")) +
  labs(x="MYA", y="log(CC)", title="Extant CC Over Time by Order")
Facet_order_plot

#Fossil CC Plot
Fossil_plot <- plot +
  geom_line(data=zachos, aes(x=age, y=rescale(filteredO18, c(-1,8))), color="blue") +
  scale_x_reverse() +
  geom_point(data=fossils , aes(x=Actual.MYA, y=log(CC), color=Order), size=4) +
  labs(x="MYA", y="log(CC)", title="Fossil CC Over Time") +
  theme(text = element_text(size=40)) +
  geom_smooth(data=extants, aes(x=Actual.MYA, y=log(CC)), color="red", method=lm, se=FALSE, inherit.aes=FALSE)
Fossil_plot

#Ancestral CCs Plot
Ancest_plot <- plot +
  geom_line(data=zachos, aes(x=age, y=rescale(filteredO18, c(-1,8))), color="blue") +
  scale_x_reverse() +
  geom_point(data=branch65, aes(x=mya, y=CC), size=4) +
  labs(x="MYA", y="log(CC)", title="Ancestral CC Over Time") +
  theme(text = element_text(size=40)) +
  geom_smooth(data=branch65, aes(x=mya, y=CC), color="red", method=lm, se=FALSE, inherit.aes=FALSE)
Ancest_plot

#Ancestral CCs Plot: w/ node labels
Ancest_plot_nodes <- plot +
  scale_x_reverse() +
  geom_point(data=branch65, aes(x=mya, y=CC), size=2) +
  geom_text(data=branch65, aes(x=mya, y=CC, label=nodes), hjust=0, vjust=0, size=4, color="purple") +
  labs(x="MYA", y="Ancestral CC", title="Ancestral CC Over Time") +
  theme(text = element_text(size=40)) +
  geom_smooth(data=branch65, aes(x=mya, y=CC), color="red", method=lm, se=FALSE, inherit.aes=FALSE)
Ancest_plot_nodes


```

## CC vs. Time (OLS)

```{r}
library(dplyr)

CCvsTime  <- function(order){
  filteredDF <- filter(fossils, Order == order)
  return(list(order=order,
              model = lm(log(CC) ~ MYA, data = filteredDF)
              ))
}

CCvsTime_results <- lapply(unique(fossils$Order), FUN=CCvsTime)

```

### CC vs. Time: Results

```{r}
invisible(lapply(CCvsTime_results, FUN = function(theList) {
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(paste(theList$order, "CC vs. Time by Order results"))
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(summary(theList$model))
}))

```

# Multiple regressions on fossils by order

## Multiple regressions with just fossils 

## 5 Ma time bins

```{r}
library(dplyr)
doSparseMod5Mya <- function(order){
  filteredDF <- filter(fossils, Order == order)
  return(list(order=order,
              model = lm(log(CC) ~ X5Mya_M_oxy + slope5MY + X5Mya_SD_oxy, data = filteredDF)
              ))
}

sparse_results_5Mya <- lapply(unique(fossils$Order), FUN=doSparseMod5Mya)

```

### Sparse model 5Mya results by order

```{r}
invisible(lapply(sparse_results_5Mya, FUN = function(theList) {
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(paste(theList$order, "results, 5Mya full model"))
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(summary(theList$model))
}))

plot <- ggplot(data=brains, aes(color=Order))

PLOTsparse_results_5Mya <- ggplot() +
  geom_point(data=fossils, aes(x=fossils$MYA, y=log(fossils$CC), color=fossils$Order)) +
  geom_abline(intercept=sparse_results_5Mya$model$coefficients[1], slope=sparse_results_5Mya$model$coefficients[2], color="blue") +
  geom_abline(intercept=sparse_results_5Mya$model$coefficients[1], slope=sparse_results_5Mya$model$coefficients[3], color="red") + 
    geom_abline(intercept=sparse_results_5Mya$model$coefficients[1], slope=sparse_results_5Mya$model$coefficients[4], color="green") 
PLOTsparse_results_5Mya
              
```

# Next the 1Mya results

```{r}
doSparseMod1Mya <- function(order){
  filteredDF <- filter(fossils, Order == order)
  return(list(order=order,
              model = lm(log(CC) ~ X1Mya_M_oxy + slope1MY + X1Mya_SD_oxy, data = filteredDF)
              ))
}

sparse_results_1Mya <- lapply(unique(fossils$Order), FUN=doSparseMod1Mya)

```

### Sparse model 1Mya results by order

```{r}
invisible(lapply(sparse_results_1Mya, FUN = function(theList) {
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(paste(theList$order, "results, 1Mya full model"))
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(summary(theList$model))
}))
```

#400K time bins

```{r}
doSparseMod400k <- function(order){
  filteredDF <- filter(fossils, Order == order)
  return(list(order=order,
              model = lm(log(CC) ~ X400k_M_oxy + slope400K + X400k_SD_oxy, data = filteredDF)
              ))
}

sparse_results_400k <- lapply(unique(fossils$Order), FUN=doSparseMod400k)
```

### Sparse model 400k results by order

```{r}
invisible(lapply(sparse_results_400k, FUN = function(theList) {
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(paste(theList$order, "results, 400k sparse model"))
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(summary(theList$model))
}))
```

#200k time bins

```{r}
doSparseMod200k <- function(order){
  filteredDF <- filter(fossils, Order == order)
  return(list(order=order,
              model = lm(log(CC) ~ X200k_M_oxy + slope200K + X200k_SD_oxy, data = filteredDF)
              ))
}

sparse_results_200k <- lapply(unique(fossils$Order), FUN=doSparseMod200k)
```

### Sparse model 200k results by order

```{r}
invisible(lapply(sparse_results_200k, FUN = function(theList) {
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(paste(theList$order, "results, 200k sparse model"))
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(summary(theList$model))
}))

```

# REPEAT BROKEN DOWN INTO SUBORDERS

## 5Mya time bins

```{r}
doSparseMod5Mya <- function(suborder){
  filteredDF <- filter(fossils, Suborder == suborder)
  return(list(suborder=suborder,
              model = lm(log(CC) ~ X5Mya_M_oxy + slope5MY + X5Mya_SD_oxy, data = filteredDF)
              ))
}

sparse_results_5Mya <- lapply(unique(fossils$Suborder), FUN=doSparseMod5Mya)
```

### Sparse model 5Mya results by Suborder

```{r}
invisible(lapply(sparse_results_5Mya, FUN = function(theList) {
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(paste(theList$suborder, "results, 5Mya sparse model"))
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(summary(theList$model))
}))

```

# #1Mya time bins

```{r}
doSparseMod1Mya <- function(suborder){
  filteredDF <- filter(fossils, Suborder == suborder)
  return(list(suborder=suborder,
              model = lm(log(CC) ~ X1Mya_M_oxy + slope1MY + X1Mya_SD_oxy, data = filteredDF)
              ))
}

sparse_results_1Mya <- lapply(unique(fossils$Suborder), FUN=doSparseMod1Mya)
```

### Sparse model 1Mya results by Suborder

```{r}
invisible(lapply(sparse_results_1Mya, FUN = function(theList) {
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(paste(theList$suborder, "results, 1Mya sparse model"))
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(summary(theList$model))
}))

```

#400k time bins

```{r}
doSparseMod400k <- function(suborder){
  filteredDF <- filter(fossils, Suborder == suborder)
  return(list(suborder=suborder,
              model = lm(log(CC) ~ X400k_M_oxy + slope400K + X400k_SD_oxy, data = filteredDF)
              ))
}

sparse_results_400k <- lapply(unique(fossils$Suborder), FUN=doSparseMod400k)
```

### Sparse model 400k results by Suborder

```{r}
invisible(lapply(sparse_results_400k, FUN = function(theList) {
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(paste(theList$suborder, "results, 400k sparse model"))
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(summary(theList$model))
}))

```

#200k time bins
```{r}
doSparseMod200k <- function(suborder){
  filteredDF <- filter(fossils, Suborder == suborder)
  return(list(suborder=suborder,
              model = lm(log(CC) ~ X200k_M_oxy + slope200K + X200k_SD_oxy, data = filteredDF)
              ))
}

sparse_results_200k <- lapply(unique(fossils$Suborder), FUN=doSparseMod200k)
```

### Sparse model 200k results by Suborder

```{r}
invisible(lapply(sparse_results_200k, FUN = function(theList) {
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(paste(theList$suborder, "results, 200k sparse model"))
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(summary(theList$model))
}))

```

# 100k time bins

```{r}
doSparseMod100k <- function(suborder){
  filteredDF <- filter(fossils, Suborder == suborder)
  return(list(suborder=suborder,
              model = lm(log(CC) ~ X100k_M_oxy + slope100K + X100k_SD_oxy, data = filteredDF)
              ))
}

sparse_results_100k <- lapply(unique(fossils$Suborder), FUN=doSparseMod100k)
```

### Sparse model 100k results by Suborder

```{r}
invisible(lapply(sparse_results_100k, FUN = function(theList) {
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(paste(theList$suborder, "results, 100k sparse model"))
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(summary(theList$model))
}))

```
## ^^^^^^^^^
## Comparisons to Hominins over the last 5MY
## ^^^^^^^^^

```{r}
library(dplyr)
#Filter out CCs from only the last 5Mya
short_fossils <- filter(fossils, MYA >= -5, Order != "Perissodactyla") %>% droplevels()

short_plot <- ggplot() + 
  geom_point(data=short_fossils, aes(x = MYA, y=log(CC), color=Order), size=4) + 
geom_line(data=zachos,aes(x=age*-1, y=rescale(filteredO18*-1,c(-1,8)), color="δ18O")) +
  xlim(-5, 0) + theme(text = element_text(size=40))
short_plot

```

# Next the 1Mya results

```{r}
SHORTdoSparseMod1Mya <- function(order){
  filteredDF <- filter(short_fossils, Order == order)
  return(list(order=order,
              model = lm(log(CC) ~ X1Mya_M_oxy + slope1MY + X1Mya_SD_oxy, data = filteredDF)
              ))
}

SHORTsparse_results_1Mya <- lapply(unique(short_fossils$Order), FUN=SHORTdoSparseMod1Mya)

```

### Last 5Mya only: 1My interval

```{r}
invisible(lapply(SHORTsparse_results_1Mya, FUN = function(theList) {
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(paste(theList$order, "results, 1Mya sparse model"))
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(summary(theList$model))
}))
```

#400K time bins

```{r}
SHORTdoSparseMod400k <- function(order){
  filteredDF <- filter(short_fossils, Order == order)
  return(list(order=order,
              model = lm(log(CC) ~ X400k_M_oxy + slope400K + X400k_SD_oxy, data = filteredDF)
              ))
}

SHORTsparse_results_400k <- lapply(unique(short_fossils$Order), FUN=SHORTdoSparseMod400k)
```

### Last 5Mya only: 400k interval

```{r}
invisible(lapply(SHORTsparse_results_400k, FUN = function(theList) {
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(paste(theList$order, "results, 400k sparse model"))
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(summary(theList$model))
}))
```

#200k time bins

```{r}
SHORTdoSparseMod200k <- function(order){
  filteredDF <- filter(short_fossils, Order == order)
  return(list(order=order,
              model = lm(log(CC) ~ X200k_M_oxy + slope200K + X200k_SD_oxy, data = filteredDF)
              ))
}

SHORTsparse_results_200k <- lapply(unique(short_fossils$Order), FUN=SHORTdoSparseMod200k)
```

### Last 5Mya only: 200k interval

```{r}
invisible(lapply(SHORTsparse_results_200k, FUN = function(theList) {
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(paste(theList$order, "results, 200k sparse model"))
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(summary(theList$model))
}))

```

#100k time bins

```{r}
SHORTdoSparseMod100k <- function(order){
  filteredDF <- filter(short_fossils, Order == order)
  return(list(order=order,
              model = lm(log(CC) ~ X100k_M_oxy + slope100K + X100k_SD_oxy, data = filteredDF)
              ))
}

SHORTsparse_results_100k <- lapply(unique(short_fossils$Order), FUN=SHORTdoSparseMod100k)
```

### Last 5Mya only: 100k interval

```{r}
invisible(lapply(SHORTsparse_results_100k, FUN = function(theList) {
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(paste(theList$order, "results, 100k sparse model"))
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(summary(theList$model))
}))

```
  