# New analysisis with Shultz and Dunbar mammals plus Boddy Primates

### Load datasets and clean up species columns and column names
```{r}
setwd("/Users/schilder/Desktop/Rdata")
brains <- read.table("brainclimatedata2.txt", header=TRUE, sep="\t")
brains$Species<- gsub(" ", "_", brains$Species)

boddyprimates <- read.table("Hominin_climatebrains_means2.csv", header=TRUE, sep=",")
boddyprimates$Species <- gsub(" ", "_", boddyprimates$Species)
names(boddyprimates) <- c("Species", "CC")
```

### Remove all Shultz and Dunbar extant primates
```{r}
primatesToDrop <- which(brains$Order == "Primates" & brains$Fossil == "No")
brains <- brains[-primatesToDrop,]
rm(primatesToDrop)
```

### Add in Boddy Extant Primates
```{r}
library(plyr) #note this is a different library than dplyr, needed for rbind.fill function
boddyprimates$Order <- "Primates"
boddyprimates$Fossil <- "No"
brains <- rbind.fill(brains, boddyprimates)
rm(boddyprimates)
```

## Redo ancestral states with composite dataset (Shultz and Dunbar with Boddy's extant primates)
```{r cache=TRUE}
library(ape)
library(geiger)
treez <- read.nexus("Bininda_Emonds_correctedDates_FULLPHYLOG.txt")
tree <- treez[[grep(pattern = "bestDates", names(treez))]] #get the best dates tree
rm(treez)

# fix typo in species name
brains[grep("pennsylvanicus", brains$Species),"Species"] <- "Microtus_pennsylvanicus"

name_matchup <- name.check(phy = tree, data = brains, data.names = brains$Species)
tree <- drop.tip(phy = tree, name_matchup$tree_not_data)
tree <- multi2di(tree)
tree$edge.length[tree$edge.length == 0] <- 0.1
brains <- brains[brains$Species %in% tree$tip.label,]

stopifnot(nrow(brains)==length(tree$tip.label))



brains_CC <- brains$CC
names(brains_CC) <- brains$Species


ancest <- ace(x = brains_CC, phy = tree, type="continuous", method="REML")
```

## Export plot (otherwise it it too big to see) file is "compositeShultzBoddyAncestStatesRecon.pdf"
```{r}
#pdf("compositeShultzBoddyAncestStatesRecon.pdf", width=20, height=200)
#plot(tree)
#nodelabels(round(ancest$ace, 2), cex=1.2)
#dev.off()
```

# Prep fossil data
```{r}
library(dplyr)
fossils <- read.table("brainclimatedata.txt", header=TRUE, sep="\t")
fossils$Species<- gsub(" ", "_", fossils$Species)
fossils <- tbl_df(fossils) %>%
  filter(Fossil=="Yes")


hominin <- read.table("Hominin_climatebrains_means.csv", header=TRUE, sep=",")
hominin$Species <- gsub(" ", "_", hominin$Species)

#combine hominins and other fossils
fossils <- rbind.fill(hominin, fossils)
```

# Multiple regressions on fossils by order

## Multiple regressions with just fossils 

## 5 Ma time bins

First I did a multiple regression of log(CC) by mean_oxy, sd_oxy, and slope_oxy, with all possible interaction terms.  This is the 'fullMod_5MA`.

Second, I did a sparse model, with no interaction terms.  This is the `sparseMod_5MA`

```{r}
doFullMod5Mya <- function(order){
  filteredDF <- filter(fossils, Order == order)
  return(list(order=order,
              model = lm(log(CC) ~ X5Mya_M_oxy * X5Mya_slope_oxy * X5Mya_SD_oxy, data = filteredDF)
              ))
}

doSparseMod5Mya <- function(order){
  filteredDF <- filter(fossils, Order == order)
  return(list(order=order,
              model = lm(log(CC) ~ X5Mya_M_oxy + X5Mya_slope_oxy + X5Mya_SD_oxy, data = filteredDF)
              ))
}



full_results_5Mya <- lapply(unique(fossils$Order), FUN=doFullMod5Mya)
sparse_results_5Mya <- lapply(unique(fossils$Order), FUN=doSparseMod5Mya)

```

### Full model 5Mya results by order
```{r}
invisible(lapply(full_results_5Mya, FUN = function(theList) {
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(paste(theList$order, "results, 5Mya full model"))
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(summary(theList$model))
}))
```

### Sparse model 5Mya results by order
```{r}
invisible(lapply(sparse_results_5Mya, FUN = function(theList) {
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(paste(theList$order, "results, 5Mya full model"))
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(summary(theList$model))
}))
```

# Next the 1Mya results
```{r}
doFullMod1Mya <- function(order){
  filteredDF <- filter(fossils, Order == order)
  return(list(order=order,
              model = lm(log(CC) ~ X1Mya_M_oxy * X1Mya_slope_oxy * X1Mya_SD_oxy, data = filteredDF)
              ))
}

doSparseMod1Mya <- function(order){
  filteredDF <- filter(fossils, Order == order)
  return(list(order=order,
              model = lm(log(CC) ~ X1Mya_M_oxy + X1Mya_slope_oxy + X1Mya_SD_oxy, data = filteredDF)
              ))
}



full_results_1Mya <- lapply(unique(fossils$Order), FUN=doFullMod1Mya)
sparse_results_1Mya <- lapply(unique(fossils$Order), FUN=doSparseMod1Mya)

```

### Full model 1Mya results by order
```{r}
invisible(lapply(full_results_1Mya, FUN = function(theList) {
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(paste(theList$order, "results, 1Mya full model"))
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(summary(theList$model))
}))
```

### Sparse model 1Mya results by order
```{r}
invisible(lapply(sparse_results_1Mya, FUN = function(theList) {
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(paste(theList$order, "results, 1Mya sparse model"))
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(summary(theList$model))
}))
```

#400K time bins
```{r}
doFullMod400k <- function(order){
  filteredDF <- filter(fossils, Order == order)
  return(list(order=order,
              model = lm(log(CC) ~ X400k_M_oxy * X400k_slope_oxy * X400k_SD_oxy, data = filteredDF)
              ))
}

doSparseMod400k <- function(order){
  filteredDF <- filter(fossils, Order == order)
  return(list(order=order,
              model = lm(log(CC) ~ X400k_M_oxy + X400k_slope_oxy + X400k_SD_oxy, data = filteredDF)
              ))
}


full_results_400k <- lapply(unique(fossils$Order), FUN=doFullMod400k)
sparse_results_400k <- lapply(unique(fossils$Order), FUN=doSparseMod400k)
```

### Full model 400k results by order
```{r}
invisible(lapply(full_results_400k, FUN = function(theList) {
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(paste(theList$order, "results, 400k full model"))
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(summary(theList$model))
}))
```

### Sparse model 400k results by order
```{r}
invisible(lapply(sparse_results_400k, FUN = function(theList) {
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(paste(theList$order, "results, 400k sparse model"))
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("||||||||||||||||||||||||")
  print("vvvvvvvvvvvvvvvvvvvvvvvv")
  print(summary(theList$model))
}))
```
